### 2014-10-03 paper/include/background.tex
modified bg

DRAM is hierarichally organized as \reffig{fig:dram_orga} shows. At the top of the hierarchy are the channels that are connected to one or more ranks. Each channel has individual address, data, and command buses that makes it possible for the channels to operate cuncurrently. All ranks on one channel can operate in parallel constrained by the channel bandwith which is shared among the ranks. A rank comprizes of several banks where one bank holds a DRAM cell array. The banks are thus also able to work in parallel, but are constrianed by both the shared channel bandwith and resources shared among the banks.

The DRAM cell array in each bank is structured as \reffig{fig:dram_orgb} shows. One cell consists of a transistor and a capacitor, where the transistor is controlled by the wordline wire and connects the capacitor to the bitline wire. The wordline connects a number of cells which forms a row. Each bitline is connected to a sens amplifier and the row of sens amplifier constitutes the row buffer. Data is stored as a charge on the cell capacitor.

@@

To read or write to a DRAM cell the wordline connected to the targeted transistor must first be activated. When the wordline is activated, the transistor connects the capacitor to the bitline. For this to work, the bitlines has to be precharged to \(V_{DD}\)/2. When the capacitor connects to the bitline, charge will either flow from or to the capacitor depending on if the capacitor was charged or not, respectively. The sens amplifier will detect the change of voltage and drive the bitline either to \(V_{DD}\) or 0 V. Data can then be read from the bitline. If data is to be written the corresponding voltage is applied to the bitline, which stores the data in the capacitor. Finally, if another row on the bank chall be accessed the open row has to be closed, which disconnects the capacitors from the bitlines.

>>>>

DRAM is hierarichally organized as in \reffig{fig:dram_orga}. At the top of the hierarchy are the channels that are connected to one or more ranks, each composed of multiple storage arrays called banks. Each channel has individual address, data, and command buses that makes it possible for the channels to operate cuncurrently. All ranks on one channel can also operate in parallel constrained by the channel bandwith which is shared among the ranks. Further, the banks within the ranks are also able to operate in parallel, but are constrianed by both the shared channel bandwith and resources shared among the banks.

The DRAM cell array in each bank is structured as in \reffig{fig:dram_orgb}. One cell consists of a transistor and a capacitor, where the transistor is controlled by the wordline wire and connects the capacitor to the bitline wire. The wordline connects a number of cells which forms a row. Each bitline is connected to a sens amplifier and the row of sense amplifier constitutes the row buffer. Data is stored as a charge on the cell capacitor.

@@

To read a row the wordline connected to the rows transistors must first be activated. When the wordline is activated, the transistors connects the capacitors to the bitlines. When the capacitors connects to the bitlines, charge will either flow from or to the capacitors depending on if the capacitors were charged or not, respectively. For this to work, the bitlines has to be precharged to \(V_{DD}\)/2. The sense amplifiers then detect the change of voltage, which makes it possible to interpret the stored data. The charge stored on the capacitors has, as mentioned, left the capacitor and a read operation is thus destructive. To maintain the data the capacitors are recarged by the bitlines, which are driven to either either \(V_{DD}\) or 0 V by the sense amplifiers depending on the former detected voltage change. Finally, if another row on the bank shall be accessed the open row has to be closed, which disconnects the capacitors from the bitlines.

To write to a row the corresponding wordline is first activated. As during a read, the transistors then connects the capacitors to the bitlines and charge will either flow from or to the capacitor. The row buffer are then loaded with the data that are driven to the bitlines by the sense amplifiers, thus storing the information on the capacitors. The row then remains open, as by a read, until another row shall be accessed, then the row is closed which disconnects the capacitors from the bitlines.
